---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(rlang__backtrace_on_error = "reminder")
```

# hardhat

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/DavisVaughan/hardhat.svg?branch=master)](https://travis-ci.org/DavisVaughan/hardhat)
[![Codecov test coverage](https://codecov.io/gh/DavisVaughan/hardhat/branch/master/graph/badge.svg)](https://codecov.io/gh/DavisVaughan/hardhat?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/hardhat)](https://cran.r-project.org/package=hardhat)
<!-- badges: end -->

## Introduction

hardhat is designed to ease the creation of new modeling packages, while simultaneously promoting good R modeling package standards as laid out by the set of opinionated [Conventions for R Modeling Packages](https://tidymodels.github.io/model-implementation-principles/).

The idea is to take as much of the burden around creating a good interface off the developer as possible, and instead let them focus on writing the core implementation of the model. This benefits not only the developer, but also the user of the modeling package, as the standardization allows users to build a set of "expectations" around what any modeling function should return, and how they should interact with it.

```{r}
library(hardhat)
```


## Installation

You can install the released version of hardhat from [CRAN](https://CRAN.R-project.org) with:

``` r
# no you cannot
install.packages("hardhat")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DavisVaughan/hardhat")
```

## Example

hardhat will mainly be useful to developers, and it allows you to develop new modeling packages that implement formula, data frame, matrix, and recipes interfaces with ease.

There are a number of useful functions in hardhat, but two of the most important ones are `mold()` and `forge()`.
