---
title: "Forging data for predictions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Forging data for predictions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hardhat)
```

```{r}
iris_train <- iris[1:100,]
iris_test <- iris[101:150,]

processed <- mold(log(Sepal.Length) ~ Species + Petal.Width, iris_train)
```


### `forge()`

`forge()` takes `new_data` and applies the same preprocessing steps that happened to the data used in training the model. It is to be called from `predict()`, or potentially from a cross validation function where you will use `predict()` repeatedly to measure performance of different folds or different hyperparameters.

Say you fit a model and called `mold()` from your fitting function. When you return the model object, you should attach the `preprocessor` that you get with the output of `mold()` onto the model object (using the model constructor, `new_base_model()`, makes this easier to do). Then when you call `predict()` on that model object along with `new_data`, you should call `forge()` inside of the `predict()` method with the stored `preprocessor` and the `new_data`.

```{r}
processed_test <- forge(
  preprocessor = processed$preprocessor, 
  new_data = iris_test,
  outcomes = TRUE
)
```

`forge()` always returns a list with two things. The first is a tibble containing the preprocessed `predictors`. The second is optionally the preprocessed `outcomes` if you are performing cross validation and used a formula or recipes interface. Because we used the formula interface to generate the `processed` object, we could set `outcomes = TRUE` above to also return the processed outcome column.

```{r}
processed_test$predictors

processed_test$outcomes
```

The nice thing about `forge()` is that the `preprocessor` remembers a lot of information about what happened at fit time, and keeps the user from shooting themselves in the foot at prediction time. 

For instance, each predictor used at fit time has to have the same class at prediction time.

```{r, error=TRUE}
iris_test_bad <- iris_test

# Turning Species into a character column rather than
# a factor
iris_test_bad$Species <- as.character(iris_test_bad$Species)

forge(processed$preprocessor, iris_test_bad)
```

And each predictor column has to exist in `new_data`.

```{r, error=TRUE}
# Removing species alltogether
iris_test_bad$Species <- NULL

forge(processed$preprocessor, iris_test_bad)
```

And new levels in any of the predictors throw a warning and are coerced to `NA`.

```{r, warning=TRUE}
iris_test_bad <- iris
iris_test_bad$Species <- as.character(iris_test_bad$Species)
iris_test_bad$Species[1] <- "new_level"
iris_test_bad$Species  <- factor(iris_test_bad$Species)

levels(iris_test_bad$Species)

processed_bad_test <- forge(processed$preprocessor, iris_test_bad)

processed_bad_test$predictors
```



