---
title: "Forging data for predictions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Forging data for predictions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(rlang__backtrace_on_error = "reminder")
```

```{r setup}
library(hardhat)
```

## Introduction

The counterpart to `mold()` (which you can read all about in `vignette("mold", "hardhat")`), is `forge()`. Where `mold()` is used to preprocess your training data, `forge()` is used to preprocess new data that you are going to use to generate predictions from your model.

Like `mold()`, `forge()` is not intended to be used interactively. Instead, it should be called from the `predict()` method for your model. To learn more about using `forge()` in a modeling package, see `vignette("package", "hardhat")`. The rest of this vignette will be focused on the many features that `forge()` offers.

## Connection with `mold()`

When `mold()` is used, one of the returned objects is an `engine`. This is the key to preprocessing new data with `forge()`. For instance, assume you've called `mold()` like so:

```{r}
iris_train <- iris[1:100,]
iris_test <- iris[101:150,]
```

```{r}
iris_form <- mold(
  log(Sepal.Length) ~ Species + Petal.Width, 
  iris_train, 
  engine = default_formula_engine(indicators = FALSE)
)

formula_eng <- iris_form$engine

formula_eng
```

A formula engine is returned here, which knows about the predictors and outcomes that were used at training time, and knows that you don't want to expand `Species` into dummy variables by setting `indicators = FALSE`.

When it is time to `predict()` on new data, that data is passed on to `forge()` along with the `engine` we just created.

```{r}
forge(iris_test, formula_eng)
```

Note that in `predictors`, `Species` was not expanded because the `engine` knew about the preprocessing options that were set when `mold()` was called.

`forge()` always returns three things, and they should look familiar to you if you have used `mold()`. 

- `predictors` holds a tibble of the predictors.

- `outcomes` is returned as `NULL` by default, because most `predict()` methods assume you only have access to the new predictors. Alternatively, as you will read in a moment, this can contain a tibble of the new outcomes.

- `extras` varies per engine, but is a catch-all slot to hold the same kind of extra objects that were returned by the engine when `mold()` was called.

## Outcomes

Generally when generating predictions you only need to know about the new predictors. However, when performing resampling you will need to have the processed outcomes as well so you can compute cross validated performance statistics and decide between multiple models, or choose between hyperparameters.

You can easily request the outcomes as well with `outcomes = TRUE`. Just like with the predictors, these get processed using the same steps as done to the outcomes at fit time.

```{r}
forge(iris_test, formula_eng, outcomes = TRUE)
```

## Validation

One of the most useful things about `forge()` is its robustness against malformed new data. It isn't unreasonable to enforce that the new data a user provides at prediction time should have the same _type_ as the data used at fit time. _Type_ is defined in the [vctrs](https://vctrs.r-lib.org/articles/type-size.html) sense, and for our uses essentially means that a number of checks on the test data have to pass, including:

- The column names of the testing data and training data must be the same.

- The _type_ of each column of the testing data must be the same as the columns found in the training data. This means:

  - The classes must be the same (e.g. if it was a factor in training, it must be a factor in testing).
  
  - The attributes must be the same (e.g. the levels of the factors must also be the same).

Almost all of this validation is possible through the use of `vctrs::vec_cast()`, and is called for you by `forge()`.

### Column existance

The easiest example to demonstrate is missing columns in the testing data. `forge()` won't let you continue until all of the required predictors used at training are also present in the new data.

```{r, error=TRUE}
test_missing_column <- subset(iris_test, select = -Species)

forge(test_missing_column, formula_eng)
```

### Column types

After an initial scan for the column names is done, a deeper scan of each column is performed, checking the type of that column. For instance, what happens if the new `Species` column was a double, not a factor?

```{r, error=TRUE}
test_species_double <- iris_test
test_species_double$Species <- as.double(test_species_double$Species)

forge(test_species_double, formula_eng)
```

An error is thrown, indicating that a double can't be cast to a factor.

### Lossless conversion

The error message above suggests that in some cases you _can_ automatically cast from one type to another, and in fact that is true! Rather than being a double, what if `Species` was just a character?

```{r}
test_species_character <- iris_test
test_species_character$Species <- as.character(test_species_character$Species)

forged_char <- forge(test_species_character, formula_eng)

forged_char$predictors

class(forged_char$predictors$Species)

levels(forged_char$predictors$Species)
```

Interesting, so in this case we can actually convert to a factor, and the class and even the levels are all restored. The key here is that this was a _lossless_ conversion. We lost no information when converting the character `Species` to a factor because the unique character values were a subset of the original levels.

An example of a conversion that would be lossy is if the character `Species` column had a value that was _not_ a level in the training data.

```{r, warning=TRUE}
test_species_lossy <- iris_test
test_species_lossy$Species <- as.character(test_species_lossy$Species)
test_species_lossy$Species[2] <- "im new!"

forged_lossy <- forge(test_species_lossy, formula_eng)

forged_lossy$predictors
```

In this case:

- A lossy warning is thrown

- The `Species` column is still converted to a factor with the right levels

- The novel level is removed and its value is set to `NA`

## Recipes and `forge()`

Just like with the formula method, a recipe can be used as the preprocessor at fit and prediction time. `hardhat` handles calling `prep()`, `juice()`, and `bake()` for you at the right times. For instance, say we have a recipe that just creates dummy variables out of `Species`.

```{r, error=FALSE, warning=FALSE}
library(recipes)

rec <- recipe(Sepal.Width ~ Sepal.Length + Species, iris_train) %>%
  step_dummy(Species)

iris_recipe <- mold(rec, iris_train)

iris_recipe$predictors
```

The engine is a `recipe` engine.

```{r}
recipe_eng <- iris_recipe$engine

recipe_eng
```

When we `forge()`, we can request `outcomes` to have the predictors and outcomes separated like with the formula method.

```{r}
forge(iris_test, recipe_eng, outcomes = TRUE)
```

### A note on recipes

One complication with `recipes` is that, in the `bake()` step, the processing happens to the predictors and the outcomes all together. This means that you might run into the situation where the outcomes seem to be required to `forge()`, even if you aren't requesting them.

```{r}
rec2 <- recipe(Sepal.Width ~ Sepal.Length + Species, iris_train) %>%
  step_dummy(Species) %>%
  step_center(Sepal.Width) # Here we modify the outcome

iris_recipe2 <- mold(rec2, iris_train)

recipe_eng_log_outcome <- iris_recipe2$engine
```

If our `new_data` doesn't have the outcome, baking this recipe will fail even if we don't request that the outcomes are returned by `forge()`.

```{r, error=TRUE}
iris_test_no_outcome <- subset(iris_test, select = -Sepal.Width)

forge(iris_test_no_outcome, recipe_eng_log_outcome)
```

The way around this is to use the built-in recipe argument, `skip`, on the step containing the outcome. This skips the processing of that step at `bake()` time.

```{r}
rec3 <- recipe(Sepal.Width ~ Sepal.Length + Species, iris_train) %>%
  step_dummy(Species) %>%
  step_center(Sepal.Width, skip = TRUE)

iris_recipe3 <- mold(rec3, iris_train)

recipe_eng_skip_outcome <- iris_recipe3$engine

forge(iris_test_no_outcome, recipe_eng_skip_outcome)
```

There is a tradeoff here that you need to be aware of. 

- If you are just interested in generating predictions on completely new data, you can safely use `skip = TRUE` because you will almost never have access to the corresponding true outcomes to preprocess and compare against. 

- If you know you need to do resampling, you will likely have access to the outcomes during the resampling step so you can cross-validate the performance. In this case, you can't set `skip = TRUE` because then the outcomes won't be processed, but since you have access to them, you shouldn't need to.

For example, if we used `iris_test` with the above recipe (which has the outcome), `Sepal.Width` wouldn't get centered when `forge()` is called. But we probably would not have skipped that step if we knew that our test data would have the outcome.

```{r}
forge(iris_test, recipe_eng_skip_outcome, outcomes = TRUE)$outcomes

# Notice that the `outcome` values haven't been centered
# and are the same as before
head(iris_test$Sepal.Width)
```
